rrrsa

### 2017-10-20 paper01_peloquin_frank2016/model_walkthru/paper01_peloquin_frank2016.rmd
source code rewrite now done

---
title: "paper01 model walkthru"
subtitle: "Peloquin & Frank (2016), 'Determining the alternatives for scalar implicature'"
author: "timothy leffel"
date: "oct20/2017"
---

### Gameplan

- recap of paper 
- what the model is (broad strokes)
- what the model does 
- ...

### Recap of Paper

### Model

```{r}
lefftpack::lazy_setup()
# devtools::install_github("benpeloquin7/rrrsa")
library("rrrsa")

>>>>

---
title: "Determining the alternatives for scalar implicature"
subtitle: "Peloquin & Frank (2016)"
author: "timothy leffel"
date: "oct20/2017"
---

```{r}
### ADAPTED CODE FROM PACKAGE VIGNETTE FOR DEMO OF fullRecursion()
source("setup.r")

# m/dat_pre ~~> matrix of semantics (rows=meaning (m rows), cols=words (n cols))
# costs     ~~> ncol(m) vector of costs (default is 0 valued vector)
# priors    ~~> nrow(m) vector of priors (default is uniform)
# alpha     ~~> decision hyper-param
dat_pre <- matrix(
  nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), c("none","some","all")), 
  data=c(1.0, .00, .00, 
         .00, .00, .00, 
         .25, .25, .25, 
         .25, .00, .00, 
         .00, .00,  1)
)
dat_pre_df <- rsa_matrix_to_df(dat_pre, response_colname="semantics")

costs <- c(none=0, some=0, all=0)
priors <- rnorm(nrow(dat_pre), mean=.5, sd=.1)
alpha <- 1

# rsa.fullRecursion(m=m, costs=rep(0, ncol(m)), priors=rep(1, nrow(m)), alpha=2)
dat_post <- fullRecursion(m=dat_pre, 
                          costs=rep(0,ncol(dat_pre)), 
                          priors=rep(1,nrow(dat_pre)),
                          alpha=alpha)

knitr::kable(dat_pre, row.names=TRUE)
knitr::kable(dat_post, row.names=TRUE)

make_rsa_plot(dat_pre, dat_post)
```

```{r}
# source("setup.r")

nams <- c("exp", "scale", "stars", "speaker.p", "words", "e11", "e6")
#   1) quantityVarName :: "stars"
#   2) semanticsVarName :: "speaker.p"
#   3) itemVarName :: "words"
#   4) groupVarName :: "scale"
#   other) experiment number :: "exp"
#   other) Pragmatic judgments study 1 (e6), study 2 (e11)
sapply(rsa_data, dim)
sapply(rsa_data, function(df){ identical(names(df), nams) })

pf_data <- rsa_data[names(rsa_data)[names(rsa_data) != "frank"]]

pf_data <- lapply(seq_along(pf_data), function(idx){
  pf_data[[names(pf_data)[idx]]] %>% 
    mutate(dataset = names(pf_data)[idx]) %>% 
    mutate_if(is.factor, as.character) %>% 
    rename(study1=e6, study2=e11)
}) %>% 
  (function(df_list) do.call(rbind, df_list))

str(pf_data)
# pf_data %>% filter(words=="none") %>% View() # ***

# carve out the <none, some, all> scale (both expts 1 and 2) 
# pf_data %>% filter(scale=="some_all") %>% View()
pf3_NSA <- pf_data %>% 
  filter(dataset=="pf3") %>% 
  filter(scale=="some_all") %>% 
  select(-exp, -scale, -dataset) %>% 
  select(stars, words, study1, study2, speaker.p) %>% 
  arrange(stars, words)

# all response fields are 0 for none, across the whole scale...
# pf3_NSA %>% filter(words=="none") %>% View() # ***

pf3_NSA_study1 <- pf3_NSA %>% 
  select(stars, words, value=study1) %>% 
  tidyr::spread(words, value) %>% 
  set_rownames(.$stars) %>%  # `stars` now encoded as rownames
  select(none, some, all) %>% 
  as.matrix()

# this is the semantic representation (prob dist) of each scalar item 
input_matrix <- pf3_NSA_study1

# need to make sure length matches ncol
costs <- c(none=0, some=0, all=0)

# need to make sure length matches nrow 
priors <- rnorm(nrow(input_matrix), mean=.5, sd=.1)

# parameter that fixes "how much the cost matters" 
alpha <- 1

# calculate the posterior distributions for each scalar item 
output_matrix <- fullRecursion(
  input_matrix, costs=costs, priors=priors, alpha=alpha
)

knitr::kable(input_matrix, row.names=TRUE)
knitr::kable(output_matrix, row.names=TRUE)

make_rsa_plot(input_matrix, output_matrix)
```

```{r}
# play around with all the params!!! 
rsa.reason(m=input_matrix, 
           costs=rep(0, ncol(input_matrix)), 
           priors=rep(1, nrow(input_matrix)), 
           depth=2, 
           alpha=1, 
           usePriorEveryRecurse=FALSE)

reason(input_matrix, 
       costs=rep(0, ncol(input_matrix)), 
       priors=rep(1, nrow(input_matrix)), 
       alpha=1, 
       depth=2, 
       recycle_priors=FALSE)

rsa.fullRecursion(input_matrix, 
                  costs=rep(0, ncol(input_matrix)), 
                  priors=rep(1, nrow(input_matrix)),
                  alpha=1)
fullRecursion(input_matrix, 
              costs=rep(0, ncol(input_matrix)), 
              priors=rep(1, nrow(input_matrix)),
              alpha=1)

```{r eval=FALSE, include=FALSE}
lefftpack::lazy_setup()
# devtools::install_github("benpeloquin7/rrrsa")
library("rrrsa")

### 2017-10-20 paper01_peloquin_frank2016/paper_and_materials/rrrsa_vignette.Rmd
source code rewrite now done

## Prep data
pragmaticsTidied <- res %>%
  gather(word, pragmatics, -quantity)
semanticsTidied <- as.data.frame(m) %>%
  mutate(quantity = rownames(.)) %>%
  gather(word, semantics, c(none, some, all))
fullData <- merge(pragmaticsTidied, semanticsTidied) %>%
  gather(type, value, c(pragmatics, semantics)) %>%
  mutate(quantity = as.numeric(quantity),
         lineSize = ifelse(word == "some", 2, 1))

>>>>

## Prep data
pragmaticsTidied <- res %>%
  tidyr::gather(word, pragmatics, -quantity)
semanticsTidied <- as.data.frame(m) %>%
  mutate(quantity = rownames(.)) %>%
  tidyr::gather(word, semantics, c(none, some, all))
fullData <- merge(pragmaticsTidied, semanticsTidied) %>%
  gather(type, value, c(pragmatics, semantics)) %>%
  mutate(quantity = as.numeric(quantity),
         lineSize = ifelse(word == "some", 2, 1))
